\speculake: Hiding Malware in Speculative Execution

### 2018-07-31 paper/related-work.tex
pass on covert channels

\subsection{Cache Side Channels}

Cache side-channels have been used against a number of processors at essentially
all different caching levels---L1 data
cache~\cite{percival2005cache,zhang2012cross,osvik2006cache}, L1 instruction
cache~\cite{aciiccmez2010new}, L2
cache~\cite{ristenpart2009hey,percival2005cache},
LLC~\cite{ristenpart2009hey,liu2015last}, and branch prediction
cache~\cite{aciiccmez2007power}. These attacks have also been performed in the
cloud~\cite{ristenpart2009hey,zhang2012cross},
browser~\cite{oren2015spy,google_cache_browser}, and security critical
encryption~\cite{yarom2014recovering,tromer2010efficient} environments,
demonstrating how ubiquitous hardware side-channel exploits can be.

Cache side channels come in many
variants~\cite{neve2006refined,tromer2010efficient,yarom2014flush+,gruss2016flush+flush}.
In \speculake, we utilize the \texttt{Flush+Reload} method, where we first flush
all items from cache, wait for the speculative world to reload a value, and then
time accesses to reloading each potential value. However, we acknowledge that
the recent \texttt{Flush+Flush} variant may offer a faster
alternative~\cite{gruss2016flush+flush}.

%\begin{itemize}
%\item \texttt{Evict+Time} - An attacker with a synchronized clock evicts a series of items 
%from cache. The victim then proceeds with computation, at which point the attacker can 
%time accesses to each evicted item to identify which items a victim loaded~\cite{neve2006refined}. 
%\item \texttt{Prime+Probe} - An attaker performs a computation, then evicts specific 
%items from the cache before performing the same action again. Information about the cache
%access patterns can be gained by comparing the timimg of the two 
%computations~\cite{tromer2010efficient}.
%\item \texttt{Flush+Reload} - Two cooperating processes  make use of the \texttt{Evict+Time} 
%model to form a communication channel~\cite{yarom2014flush+}.
%\item \texttt{Flush+Flush} - Follows the same model as \texttt{Prime+Probe},
%however clflush is used instead of a load as it resolves much quicker in cases 
%where the item is not in cache~\cite{gruss2016flush+flush}.
%\end{itemize}

%While other works have demonstrated higher throughput from similar 
%covert cache side channels~\cite{liu2015last}, \speculake incurrs slightly higher performance 
%penalties in our implementation as we perform redundancy checks to 
%ensure the signal strength.  This redundancy combined with the intentional cache misseses
%which enable speculative execution cost hundreds to thousands of clock cycles
%per round.

>>>>

\subsection{Covert Channels}

There is a plethora of previous research into cache side-channels, examining how
information is leaked between
processes~\cite{percival2005cache,zhang2012cross,osvik2006cache}. This is a key
feature of Spectre~\cite{spectre} and Meltdown~\cite{meltdown}. In \speculake we
use the branch predictor side-channel as a \emph{covert
channel}~\cite{lampson1973note} between an innocent victim program and malware.
This allows \speculake to use the leaked information from cache as a message to
a colluding malware to make decisions.

Previous work has examined various ways of sharing information over covert
channels, such as across virtualized environments on cloud
systems~\cite{wu2012whispers}, using L1 and L2 cache to share
information~\cite{percival2005cache}, measuring temperature to create a thermal
covert channel~\cite{masti2015thermal, bartolini2016capacity} and taking
advantage of processor architecture to leak information~\cite{wang2006covert}.

The branch predictor has been used as a covert
channel~\cite{evtyushkin2016understanding}. Previous work has examined the
capability of two colluding programs to share sensitive information between each
other using the branch predictor by timing how long a spy process executes taken
(not-taken) branches~\cite{evtyushkin2015covert}.

\speculake takes advantage of a cache side-channel in \texttt{OpenSSL} and uses
it as a covert channel. \texttt{OpenSSL} is not a malicious colluding program in
this case but rather a program leaking information on which cipher suite has
been selected by attackers, allowing us to use the branch predictor as an
attacker controlled covert channel of an innocent victim program.

% cache~\cite{aciiccmez2010new}, L2
% cache~\cite{ristenpart2009hey,percival2005cache},
% LLC~\cite{ristenpart2009hey,liu2015last}, and branch prediction
% cache~\cite{aciiccmez2007power}. These attacks have also been performed in the
% cloud~\cite{ristenpart2009hey,zhang2012cross},
% browser~\cite{oren2015spy,google_cache_browser}, and security critical
% encryption~\cite{yarom2014recovering,tromer2010efficient} environments,
% demonstrating how ubiquitous hardware side-channel exploits can be.

% Cache side channels come in many
% variants~\cite{neve2006refined,tromer2010efficient,yarom2014flush+,gruss2016flush+flush}.
% In \speculake, we utilize the \texttt{Flush+Reload} method, where we first flush
% all items from cache, wait for the speculative world to reload a value, and then
% time accesses to reloading each potential value. However, we acknowledge that
% the recent \texttt{Flush+Flush} variant may offer a faster
% alternative~\cite{gruss2016flush+flush}.
